---
title: "CONI Diet"
output: html_document
date: '2023-04-25'
editor_options: 
  chunk_output_type: console
---

# Initialization

Load packages

```{r}
#### To install qiime2R:
# if (!requireNamespace("devtools", quietly = TRUE)){install.packages("devtools")}
# devtools::install_github("jbisanz/qiime2R")


library(qiime2R)
library(tidyverse)

```

Custom Functions
```{r}
# length of which
length.which <- function(x) {
  length(which(x))
}

# standard error
sterr <- function(x) sd(x)/sqrt(length(x))
```
# Prepare Data

Convert .qza objects to data.frames
```{r}

# feature table
feature_qza <- read_qza("../data/table-176-dada2.qza")
sequences <- as.data.frame(feature_qza[["data"]]) %>%
  rownames_to_column("Feature.ID")  %>%
  rename_with(~ gsub("-", ".", .)) # replace "-" in column names with "."
write.csv(sequences, "../output/seq_table.csv")

# taxonomy
ncbi_taxonomy_qza <- read_qza("../data/ncbi-taxonomy.qza")
ncbi_taxonomy <- ncbi_taxonomy_qza[["data"]] %>%
  mutate(source = "ncbi")

bold_taxonomy_qza <- read_qza("../data/bold-taxonomy.qza")
bold_taxonomy <- bold_taxonomy_qza[["data"]] %>%
  mutate(source = "bold")

cust_taxonomy_qza <- read_qza("../data/cust-taxonomy.qza")
cust_taxonomy <- cust_taxonomy_qza[["data"]] %>%
  mutate(source = "cust")


```


Clean up data: Delete ASVs with <90% confidence, 
```{r}

# rbind taxonomy data frames 
long_dat <- rbind(bold_taxonomy, ncbi_taxonomy, cust_taxonomy)


# filter out matches < 90% and separate taxon columns 
filtered_dat <- long_dat %>%
  filter(Confidence >= 0.9) %>% # delete rows with confidence < 0.9
  separate(col = Taxon,           # separate taxon into multiple columns
           into = c("kingdom", "phylum", "class", "order", 
                    "family", "genus", "species"),
           sep = ";") %>%
  filter(phylum == "p__Arthropoda" | phylum == " p__Arthropoda") %>% 
  # keep only Arthropoda (some have space before p)
  arrange(Feature.ID) # arrange df by Feature.ID
# Warning message ok

# merge data taxon columns with sequence data
merged_dat <- full_join(x = filtered_dat, y = sequences, by = "Feature.ID")

# create df with negatives and blanks only
neg_dat <- merged_dat %>%
  select("Feature.ID", "kingdom", "phylum", "class", "order", "family", 
         "genus", "species", "Confidence", "source",
         "CONI.BLANK.1", "CONI.BLANK.2", "CONI.neg.1", "CONI.neg.10", 
         "CONI.neg.11", "CONI.neg.12", "CONI.neg.13", "CONI.neg.14", 
         "CONI.neg.2", "CONI.neg.3", "CONI.neg.4", "CONI.neg.6", "CONI.neg.7", 
         "CONI.neg.8", "CONI.neg.9") %>% # select negatives and blanks
  filter((CONI.BLANK.1 | CONI.BLANK.2 | CONI.neg.1 | CONI.neg.10 | 
                  CONI.neg.11 | CONI.neg.12 | CONI.neg.13 | CONI.neg.14 |
                  CONI.neg.2 | CONI.neg.3 | CONI.neg.4 | CONI.neg.6 | 
                  CONI.neg.7 | CONI.neg.8 | CONI.neg.9) != 0) 
                  # get rid of rows with all 0s
  
# create df with samples only
nonneg_dat <- merged_dat %>%
  select(-c("CONI.BLANK.1", "CONI.BLANK.2", "CONI.neg.1", "CONI.neg.10", 
         "CONI.neg.11", "CONI.neg.12", "CONI.neg.13", "CONI.neg.14", 
         "CONI.neg.2", "CONI.neg.3", "CONI.neg.4", "CONI.neg.6", "CONI.neg.7", 
         "CONI.neg.8", "CONI.neg.9"))  

# remove features from sample df that appear in negative/blank df
uncontaminated_dat <- nonneg_dat[
  !(nonneg_dat$Feature.ID %in% neg_dat$Feature.ID),]
# removed 73 ASVs

# remove features that didn't have a match to Order
assigned_dat <- uncontaminated_dat %>%
  drop_na(order) %>% # remove rows that have NA in order column
   filter(!if_all(starts_with("CONI"), is.na)) # get rid of rows with all NAs


```

# Finalize classifications


Several RSM samples matched Mantidae (o=Mantodea) to a high degree in the custom reference library, but they matched different orders in BOLD/NCBI (mostly Lepidoptera). I examined the first 25 of these ASVs and the BOLD/NCBI record appears to be accurate (the record is from Argentina or nearby), so I will have BOLD/NCBI results always outweigh Mantispidae from my custom library.	This also occurred with Calymmus.						

```{r}

# remove f__Mantispidae and  g__Calymmus (TenebrionidaeA)
dat_no_calmant <- assigned_dat %>%
  filter(genus != " g__Calymmus") %>% # remove g__Calymmus
  filter(family != " f__Mantispidae") # remove f__Mantispidae

```


In the cases where an ASV matched different taxa depending on the library, I went with the hit with the highest confidence.

```{r}
# Which features matched to just one reference library?
unique_dat <- dat_no_calmant %>%
  group_by(Feature.ID) %>%
  filter(n()==1) %>% # extract Feature.IDs that appear only once
  ungroup()

# Which features matched to multiple reference libraries?
duplicate_dat <- dat_no_calmant %>%
  group_by(Feature.ID) %>% 
  filter(n()>1) %>% # extract Feature.IDs that appear more than once
  ungroup()


# Manually remove exceptions to my highest confidence rule. 
# These exceptions were made based on geography data available in BOLD or high 
# accuracy for low taxonomic resolution.
# Then identify the hit with the highest confidence
duplicate_highest_conf <- duplicate_dat %>%
  filter(!(Feature.ID == "1aaff0383c2e616727f1d81bd7035a91" & source != "bold"),
         !(Feature.ID == "1d31a018bdea736adfa6dea80789eae7" & source != "ncbi"),
         !(Feature.ID == "608b0db3aa937e07cc16c559963b2ad1" & source != "bold"),
         !(Feature.ID == "688f6ee25d6cb45649d599a1f10e82ea" & source != "ncbi"),
         !(Feature.ID == "95d632e9afe8b913e8a6ded8eed550de" & source != "ncbi"),
         !(Feature.ID == "a263902ac5b853857b3abd5e25cbda4c" & source != "bold"),
         !(Feature.ID == "f11449ba897f6ab155bdf1a576ed2ca7" & source != "ncbi"),
         !(Feature.ID == "f8a14bf2ba4dd8099fb47e20fa8093cc" & source != "ncbi"),
         !(Feature.ID == "3d78b851e8bf227e8ef25d99f9dc2709" & source != "ncbi")
           ) %>%
  group_by(Feature.ID) %>%
  slice_max(Confidence)

# combine with unique Feature.ID df
unique_asv <- rbind(unique_dat, duplicate_highest_conf)
# confirm they're all unique
nrow(unique_asv) == length(unique(unique_asv$Feature.ID)) 

# remove the taxonomic classification codes in front of taxonomy
unique_asv$kingdom <- gsub(" ","",as.character(unique_asv$kingdom))
unique_asv$kingdom <- gsub("tax=Animalia","Metazoa",as.character(unique_asv$kingdom))
unique_asv$kingdom <- gsub("Animalia","Metazoa",as.character(unique_asv$kingdom))

unique_asv$phylum <- gsub("p__","",as.character(unique_asv$phylum))
unique_asv$phylum <- gsub(" ","",as.character(unique_asv$phylum))

unique_asv$class <- gsub("c__","",as.character(unique_asv$class))
unique_asv$class <- gsub(" ","",as.character(unique_asv$class))

unique_asv$order <- gsub("o__","",as.character(unique_asv$order))
unique_asv$order <- gsub(" ","",as.character(unique_asv$order))

unique_asv$family <- gsub("f__","",as.character(unique_asv$family))
unique_asv$family <- gsub(" ","",as.character(unique_asv$family))

unique_asv$genus <- gsub("g__","",as.character(unique_asv$genus))
unique_asv$genus <- gsub(" ","",as.character(unique_asv$genus))

unique_asv$species <- gsub("s__","",as.character(unique_asv$species))

```

Now classify species based on Zeale, Evens et al. 2020 and Jusino 2019.
I chose an approach that is more conservative than Jusino but less conservative than Evens (e.g., Jusino says >97% can be species, whereas Evens says >99 can be species, >98% genus, and anything <90% thrown away. )

```{r}

# Assign taxonomic identity based on threshold
asv_filtered <- unique_asv %>% 
  mutate(identity = ifelse(Confidence >= 0.99, species,
                    ifelse(Confidence < 0.99 & Confidence >= 0.97, genus,
                    ifelse(Confidence < 0.97 & Confidence >= 0.95, family,
                    order)))) %>%
  mutate(lowest_taxon = ifelse(identity == "" | is.na(identity),
                        ifelse(species != "" & species != " ", species,
                          ifelse(genus != "" & genus != " ", genus,
                            ifelse(family != "" & family != " ", family,
                              order))),
                        identity)) %>%
  select(lowest_taxon, everything(), -identity) 


```


## Inspect non-flying taxa

```{r}

arachnida <- asv_filtered %>%
  filter(class == "Arachnida") %>%
  select(1:11, starts_with("CONI.2")) %>%
  select(-where(~ all(is.na(.))))
unique(arachnida$order)

# We have order=Trombidiformes and Sarcoptiformes (mites), order=Ixodida (ticks), two unclassified arachnids, and four order=Araneae

```

## Further filter ASVs to get a final, clean data frame
```{r}
# Filter out non-insect taxa, add taxonomic level identifiers.
# Make blank cells with taxonomic IDs lower than lowest_taxon column
asv_unedited <- asv_filtered %>%
  mutate(lowest_taxon_grade = 
    ifelse(lowest_taxon == species, "species", 
           ifelse(lowest_taxon == genus, "genus",
                  ifelse(lowest_taxon == family, "family",
                         "order")))) %>%
  filter(class == "Insecta") %>%
  mutate(species2 = ifelse((lowest_taxon_grade == "genus" |
                              lowest_taxon_grade == "family" |
                              lowest_taxon_grade == "order") &
                             species != "", 
                           "", species)) %>%
  mutate(genus2 = ifelse((lowest_taxon_grade == "family" | 
                            lowest_taxon_grade == "order") & 
                           (genus != "" | species != ""), 
                         "", genus)) %>%
  mutate(family2 = ifelse(lowest_taxon_grade == "order" & 
                            (family != "" | genus != "" | species != ""),
                          "", family)) %>%
  select(Feature.ID, kingdom, phylum, class, order, family2, genus2, species2,
         lowest_taxon, lowest_taxon_grade, starts_with("CONI")) %>%
  dplyr::rename(species = species2, genus = genus2, family = family2) %>%
  select(where(~ !is.numeric(.) || sum(.) > 0)) # Remove columns with no reads

# write to .csv
write.csv(asv_unedited, "../output/asv_unedited.csv")

```

I then went through the "asv_unedited.csv" file and manually fixed errors (mispelled taxonomies, etc). The final, clean and edited version is called "asv_edited.csv".

```{r}

asv_final <- read.csv("../output/asv_edited.csv")

```


